---
title: "Reproducible Research: Peer Assessment 1"
author: "Celine B."
date: "March 08, 2020"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Steps required to load the data:
    1. Set working directory
    2. Unzip file
    3. Read csv file
    
    ```{r loading}
    setwd("/Users/cline/Desktop/Documents/Coursera/DSrepo/RepData_PeerAssessment1")
    getwd()
    list.files()
    unzip (zipfile = "activity.zip")
    list.files()
    activity <- read.csv("activity.csv", header = TRUE)
    ```
    
    
Exloring the data
    ```{r exploring}
    str(activity)
    summary(activity)
    head(activity)
    ```
    
Setups and packages that will be required
    ```{r setups}
    install.packages("dplyr")
    install.packages("ggplot2")
    Sys.setlocale("LC_TIME", "English_United States")
    ```


## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
